RNeXML: a package for reading and writing richly annotated phylogenetic, character, and trait data in R

### 2014-10-06 manuscripts/manuscript.Rmd
Edits for the abstract.
Mostly smoothing flow and enhancing clarity. No substantive changes.

---
#layout: preprint
layout: review, 11pt
title: "RNeXML: a package for reading and writing richly annotated phylogenetic, character, and trait data in R"
author: 
  - name: Carl Boettiger
    affiliation: cstar
    email: cboettig(at)gmail.com
    footnote: Corresponding author
  - name: Hilmar Lapp
    affiliation: NESCent
  - name: Scott Chamberlain
    affiliation: ropensci
  - name: Rutger Vos
    affiliation: NBC
address: 
  - code: cstar
    address: | 
      Center for Stock Assessment Research, 
      Department of Applied Math and Statistics, 
      University of California, Mail Stop SOE-2,
      Santa Cruz, CA 95064, USA
  - code: NESCent
    address: | 
      National Evolutionary Synthesis Center, Durham, NC, USA
  - code: ropensci
    address: |
      University of California, Berkeley, CA, USA
  - code: NBC
    address: | 
      Naturalis Biodiversity Center, Leiden, the Netherlands
abstract: | 
      1. NeXML is a new phylogenetic informatics data and metadata exchange 
      format that offers substantial improvements in reliability, extensibility, 
      and richness over earlier formats such as NEXUS. In this paper we present 
      the RNeXML package which provides a bridge between this rich and reliable
      data format and the extensive library of phylogenetic tools already
      available in R, with a particular emphasis on comparative methods.

      2. RNeXML is much richer than a basic parser: here we illustrate
      how the package facilitates adding and manipulating semantic 
      annotations, that is annotations whose meaning can be defined externally
      by way of semantic web technologies.

      3. An essential feature of NeXML is to provide both the flexibility
      to be continually extended to meet the emerging needs of the phylogenetics
      community while remaining backwards-compatible with software developed
      for earlier versions.  Here we illustrate how RNeXML makes it easy
      <!-- earlier versions of what? -->
      to extend NeXML in this manner.

      4. RNeXML is well suited for users to preserve and maintain their
      personal archives of phylogenetic and comparative trait data in consistent,
      query-able, and cross-platform format, to leverage the rapidly growing 
      public community archives of rich NeXML files, and to interact with web
      services that deliver their results (with their provenance) as NeXML.

>>>>

---
#layout: preprint
layout: review, 11pt
title: "RNeXML: a package for reading and writing richly annotated phylogenetic, character, and trait data in R"
author: 
  - name: Carl Boettiger
    affiliation: cstar
    email: cboettig(at)gmail.com
    footnote: Corresponding author
  - name: Hilmar Lapp
    affiliation: NESCent
  - name: Scott Chamberlain
    affiliation: ropensci
  - name: Rutger Vos
    affiliation: NBC
address: 
  - code: cstar
    address: | 
      Center for Stock Assessment Research, 
      Department of Applied Math and Statistics, 
      University of California, Mail Stop SOE-2,
      Santa Cruz, CA 95064, USA
  - code: NESCent
    address: | 
      National Evolutionary Synthesis Center, Durham, NC, USA
  - code: ropensci
    address: |
      University of California, Berkeley, CA, USA
  - code: NBC
    address: | 
      Naturalis Biodiversity Center, Leiden, the Netherlands
abstract: | 
      1. NeXML is a new exchange format for phylogenetic data and
      metadata that offers substantial improvements in reliability,
      extensibility, and richness over earlier formats such as
      NEXUS. In this paper we present the RNeXML package which
      provides a bridge between this data format and the extensive
      library of phylogenetic tools already available in the popular
      data analysis environment and programming language R, with a
      particular emphasis on comparative methods.

      2. The ability to parse and write NeXML from within R is only a
      small part of the functionality offered by the package.  We
      illustrate how RNeXML facilitates adding and manipulating data
      annotations whose meaning can be defined in a precise and
      machine-interpretable way by means of formal ontologies and other
      semantic web technologies.

      3. An essential feature of NeXML is to provide both the
      flexibility to be continually extended to meet the emerging
      needs of the phylogenetics community without breaking software
      unaware of such extensions. We illustrate how RNeXML makes
      it easy to extend NeXML in this manner.

      4. RNeXML is well suited for users to preserve and maintain
      their personal archives of phylogenetic and comparative trait
      data in consistent, query-able, and cross-platform format, to
      leverage the rapidly growing public community archives of rich
      NeXML files, and to interact with web services that return data
      and their metadata packaged together as NeXML.
